---
title: "TZ SNP Scatter"
author: "Hannah Koenker"
date: "11/11/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(janitor)
library(ggthemes)
library(cowplot)
theme_set(theme_light())
pall <-  "Paired"
```

```{r read, message = FALSE, echo = FALSE}
snp <- read_dta("snp_scatter.dta") %>% 
  clean_names() %>% 
  mutate(ucc3 = case_when(
    sinceucc<=2 ~ 0,
    sinceucc>2 ~ 1
    ))
```

Make two plots with all the data from the SNP councils (19 outliers dropped; boundary/name changes led to mismatch of population and ITNs that went in, sometimes TCs and DCs were combined in distribution records for example). Plot B drops councils for SNPs that occurred the year or second year after the mass campaign (i.e. Lake Zone got the 2015 MRC and then started a small SNP the next year, see bottom right corner of Plot A). Fit a line. 

```{r snp_plot, message = FALSE, echo = FALSE}
meth <- "loess"
pall <-  "Paired"

allsnp <- ggplot(snp, aes(x = accrk, y = percpop, color = as.factor(snp), alpha = 0.5)) +
  geom_point() +
  scale_alpha(guide = "none") +
  scale_color_brewer(palette = pall) +
  labs(color="SNP",
       title = "All SNP Councils (n=512)",
       x = "Estimated ITN access",
       y = "Nets distributed as a percentage of the population (NPP)") +
  xlim(0,100) +
  scale_x_continuous(breaks=seq(0, 100, 20), limits=c(0, 100))
# +
#   geom_smooth(data = snp, aes(x = accrk, y = percpop), se = FALSE, inherit.aes = FALSE, method = meth, color = "steelblue") 
ggsave("snp_fit_scatter.png")

test <- snp %>% 
  filter(ucc3==1) 

snp3 <- snp %>% 
  filter(ucc3==1) %>% 
  ggplot(aes(x = accrk, y = percpop, color = as.factor(snp), alpha = 0.5)) +
  geom_point() +
  scale_alpha(guide = "none") +
  scale_color_brewer(palette = pall) +
  labs(color="SNP",
       title = "SNP Councils 3+ years post-campaign (n=416)",
       x = "Estimated ITN access",
       y = "")  +
  xlim(0,100) +
  scale_x_continuous(breaks=seq(0, 100, 20), limits=c(0, 100)) +
  geom_smooth(data = subset(snp, ucc3==1), aes(x = accrk, y = percpop), se = FALSE, inherit.aes = FALSE, method = meth, color = "steelblue") +
  theme(legend.position = "none")
ggsave("snp3_fit_scatter.png")

plot_grid(allsnp, snp3, labels = "AUTO")
ggsave("snp_bifit_scatter.png")

```
