---
title: "05 Check Annual vs Quarterly Estimates"
author: "Hannah Koenker"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(haven)
library(janitor)
```

## Checking annual vs quarterly estimates of ITN access

Quarterly estimates are slightly higher than the annual estimates when looking at the excel output files. There may be a problem with the way that I converted the decay equation from annual to quarterly. 

Let's look first at one council to compare the trend lines. Here's Biharamulo DC. Quarterly dates are 2021.25, 2021.50, 2021.75, and 2022, for 2021q1, 2021q2, 2021q3, 2021q4, respectively. Annual dates are shifted so that access is shown at 2021.9, 2022.9, etc.

Regardless of the alignment approach (which I'm not wedded to), annual estimates seem to decay faster than the quarterly estimates.


```{r read}
qb <- read_dta("quarterly_council_itns_database.dta") %>% 
  clean_names() %>% 
  filter(council=="Biharamulo DC") 

ab <- read_dta("council_itns_access_database.dta") %>% 
  clean_names() %>% 
  filter(council=="Biharamulo DC")

qk <- read_dta("quarterly_council_itns_database.dta") %>% 
  clean_names() %>% 
  filter(council=="Kigoma DC") 

ak <- read_dta("council_itns_access_database.dta") %>% 
  clean_names() %>% 
  filter(council=="Kigoma DC")
```


```{r plotB}
ggplot() +
  geom_line(data=ab, aes(x=year+0.5, y=access_est, color="annual")) +
  geom_line(data=qb, aes(x=qyear, y=access_est, color="quarterly")) +
  theme_minimal() +
  scale_x_continuous(breaks=seq(2004,2030,1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1, size=6)) +
  labs(x="",
       y="Estimated ITN access",
       color="",
       title="Biharamulo DC")
```

```{r plotK}
ggplot() +
  geom_line(data=ak, aes(x=year+.5, y=access_est, color="annual")) +
  geom_line(data=qk, aes(x=qyear, y=access_est, color="quarterly")) +
  theme_minimal() +
  scale_x_continuous(breaks=seq(2004,2030,1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1, size=6)) +
  labs(x="",
       y="Estimated ITN access",
       color="",
       title="Kigoma DC")
```

```{r plotK-nets, include=FALSE}
ggplot() +
  geom_line(data=ak, aes(x=year+.5, y=totalnets, color="annual")) +
  geom_line(data=qk, aes(x=qyear, y=totalqnets, color="quarterly")) +
  theme_minimal() +
  scale_x_continuous(breaks=seq(2004,2030,1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1, size=6)) +
  labs(x="",
       y="NPP",
       color="",
       title="Kigoma DC")
```

In my code, annual net crops are generated with a 4-stage 'if' function, borrowed from Amelia, representing 4 possible situations - either the current year/quarter is a distribution so there's no decay, or the current year/quarter is before a distribution and is left as zero, or the current year/quarter is so far after a distribution that all the nets are gone, or it's after a distribution but before all nets are gone, in which case the decayed crop is calculated with:

``` {r stata-function, echo=TRUE, eval=FALSE}
replace cropp`j'=cropp`j'[_n-`tdist']*exp(k-(k/(1-((`tdist'/L)*(`tdist'/L))))) if cropp`j'[_n-`tdist']==totalnets[_n-`tdist'] & cropp`j'[_n-`tdist']!=. & totalnets[_n-`tdist']!=0 & cropn==`i'
```

*Parameters:*

* k = 20 - this is a constant

* L = time in years at which all nets are gone/decayed. In the annual function this is 11.75, which is the value for a median loss time of 2.15. In the quarterly function I solved for L by setting the median loss time to 2.15x4 quarters, or 8.60, and then solved for the L value in quarters, which was 70.62.

*BUT MAYBE THAT APPROACH IS NOT CORRECT.*

definitions of locals used below:

* j - numbering of the years since the nets were distributed, 1-27, used for numbering crop variables (crop1, crop2 etc)

* i - numbering of the rows (years), to calculate time since distribution and calculate decay based on the original distributed nets

* tdist = time (years) since distribution

Other:

* in Stata, [n-1] after a variable means you're referencing the value in the row before 


``` {r stata-annual, echo=TRUE, eval=FALSE}
foreach j of numlist 1/27 {
			foreach i of numlist 1/27 {
				local tdist=`i'-`j'
				if `i'==`j' {
					// do nothing, it's a distribution, leave net quantity in place
				}
				if `i'<`j' {
					// tdist is a negative number, is before distribution, leave it as missing, will replace with 0 later
				}
				if `tdist'>L {
					// e.g. if 13-7 is more than 11.75, make crop2=0
				}
				if `tdist'>0 & `tdist'<L {
					replace cropp`j'=cropp`j'[_n-`tdist']*exp(k-(k/(1-((`tdist'/L)*(`tdist'/L))))) if cropp`j'[_n-`tdist']==totalnets[_n-`tdist'] & cropp`j'[_n-`tdist']!=. & totalnets[_n-`tdist']!=0 & cropn==`i'
				}
			} // go to next i (years since distribution)
		} // go to next j (crop column)
```		
		
		
In the Quarterly calculations, the code is very similar, except we have 108 quarters instead of 27 years of 'time' in the dataset, and we're using the quarterly L values:

``` {r stata-quarterly, echo=TRUE, eval=FALSE}
foreach j of numlist 1/108 {
			foreach i of numlist 1/108 {
				local tdist=`i'-`j'
				if `i'==`j' {
					// do nothing, it's a distribution, leave net quantity in place
				}
				if `i'<`j' {
					// tdist is a negative number, is before distribution, leave it as missing, will replace with 0 later
				}
				if `tdist'>L {
				*	replace crop`j'=0 if crop`j'[_n-`i']==totalnets[_n-`i'] & crop`j'[_n-`i']!=.
				* replace crop`j'=0 if crop`j'[_n-`i']!=totalnets[_n-`i'] & crop`j'[_n-`i']!=.
					* if 80-5 is more than 70.62 (too many quarters have passed), all nets are gone, so leave this crop = 0
				}
				if `tdist'>0 & `tdist'<L {
					replace cropq`j'=cropq`j'[_n-`tdist']*exp(k-(k/(1-((`tdist'/L)*(`tdist'/L))))) if cropq`j'[_n-`tdist']==totalqnets[_n-`tdist'] & cropq`j'[_n-`tdist']!=. & totalqnets[_n-`tdist']!=0 & cropqn==`i'
				}
			} // go to next i (quarters since distribution)
		} // go to next j (crop column)
```

```{r}
small <- qk %>% 
filter(year==2022 | year==2023) %>% 
select(council, year, quarter, totalqnets, access_est)
```

