---
title: "06_Councils_QvsA_Loop"
author: "Hannah Koenker"
date: "`r Sys.Date()`"
output: pdf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(haven)
library(janitor)
```

## Checking annual vs quarterly estimates of ITN access

This file shows graphs of each council using quarterly timesteps (e.g. 2015.0, 2015.25, 2015.50, 2015.75) against the annual timesteps (mid year, so 2015.50) for ITN access estimates.

I set up the counters to track the loops (and create new crop variables) separately from time-since-distribution (tdist) in the Stata code; tdist was divided by 4 but counters were set for 27 years in the annual distribution, and for 108 quarters in the quarterly distribution. 

I think this is about as aligned as we can get. Quarterly access shows drops that are a bit masked by the annual numbers. 


```{r read}
q <- read_dta("quarterly_council_itns_database.dta") %>% 
  clean_names() %>% 
  mutate(yearq=year+(quarter/4))

a <- read_dta("council_itns_access_database.dta") %>% 
  clean_names() 


```


```{r plotloop}
for ( c in unique(a$council) ) {
  aa <- a %>% filter(council==c)
  qq <- q %>% filter(council==c)
  
  p <- ggplot() +
  geom_line(data=aa, aes(x=year+.5, y=access_est, color="annual")) +
  geom_line(data=qq, aes(x=qyear, y=access_est, color="quarterly")) +
  theme_minimal() +
  scale_x_continuous(breaks=seq(2004,2030,1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1, size=6)) +
  labs(x="",
       y="Estimated ITN access",
       color="",
       title=as.character(c))
  print(p)
}

```
