---
title: "TZ SNP Scatter"
author: "Hannah Koenker"
date: "11/11/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(janitor)
library(ggthemes)
library(cowplot)
library(dplyr)
theme_set(theme_light())
pall <-  "Paired"
```

```{r read, message = FALSE, echo = FALSE}
snp <- read_dta("snp_scatter.dta") %>% 
  clean_names() %>% 
  mutate(ucc3 = case_when(
    sinceucc<=2 ~ 0,
    sinceucc>2 ~ 1
    ),
    label = as.factor(paste0("SNP ",snp)))

## 578 SNP council-years
```

Make two plots with all the data from the SNP councils (19 outliers dropped; boundary/name changes led to mismatch of population and ITNs that went in, sometimes TCs and DCs were combined in distribution records for example). Plot B drops councils for SNPs that occurred the year or second year after the mass campaign (i.e. Lake Zone got the 2015 MRC and then started a small SNP the next year, see bottom right corner of Plot A). Fit a line. 

```{r snp_plot_predicted, message = FALSE, echo = FALSE}
# meth <- "gam"
pall <-  "Paired"

allsnp <- snp %>% 
  filter(outlier==0) %>% 
  ggplot(aes(x = access_est, y = percpop, color = as.factor(snp), alpha = 0.5)) +
  geom_point(shape=20) +
  scale_alpha(guide = "none") +
  scale_color_brewer(palette = pall) +
  labs(color="SNP",
       title = "All SNP Councils (n=512)",
       x = "Predicted ITN access",
       y = "Nets issued as a percentage of the population (NPP)") +
  xlim(0,100) +
     scale_y_continuous(breaks=seq(0, 40, 10), limits=c(0, 40)) +
scale_x_continuous(breaks=seq(0, 100, 20), limits=c(0, 100)) +
   theme(plot.title = element_text(size = 9)) +
   geom_curve(curvature=0.5, arrow = arrow(length = unit(0.03,"npc")), lineend = "round", x=75, xend=85, y=0.5, yend=3, color="dimgrey") +
  annotate("text", x=45, y=0, label="Small SNP distributions following 2015 MRC", size=2)
allsnp
# +
#   geom_smooth(data = snp, aes(x = accrk, y = percpop), se = FALSE, inherit.aes = FALSE, method = meth, color = "steelblue") 
ggsave("snp_fit_scatter.png")


snp3 <- snp %>% 
  filter(ucc3==1 & outlier==0) %>% 
  ggplot(aes(x = access_est, y = percpop, color = as.factor(snp), alpha = 0.5)) +
  geom_point(shape=20) +
  scale_alpha(guide = "none") +
  scale_color_brewer(palette = pall) +
  labs(color="SNP",
       title = "SNP Councils ≥3 years post-campaign (n=416)",
       x = "Predicted ITN access",
       y = "")  +
  xlim(0,100) +
   scale_y_continuous(breaks=seq(0, 40, 10), limits=c(0, 40)) +
 scale_x_continuous(breaks=seq(0, 100, 20), limits=c(0, 100)) +
  geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & access_est>40), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "loess", span=2, color = "steelblue") +
  # geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & access_est>40), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "glm", color = "red") +
  theme(legend.position = "none",
        plot.title = element_text(size = 9))
ggsave("snp3_fit_scatter.png")

plot_grid(allsnp, snp3, labels = "AUTO", axis="l", align="hv")
ggsave("snp_bifit_scatter.png", width=7, height=5, units="in")

```

```{r snp_plot_surveys, message = FALSE, echo = FALSE}
# meth <- "gam"
pall <-  "Paired"

allsnpsurv <- snp %>% 
  filter(percpop<40) %>% 
  ggplot(aes(xmin = access_survlb, xmax=access_survub, y = percpop, color = as.factor(snp), alpha = 0.5)) +
  geom_linerange() +
  geom_point(aes(x=access_surv, y=percpop, color=as.factor(snp), alpha=0.5), shape=10) +
  scale_alpha(guide = "none") +
  scale_color_brewer(palette = pall) +
  labs(color="SNP",
       title = "All SNP Councils (n=456)",
       x = "Survey ITN access",
       y = "Nets distributed as a percentage of the population (NPP)") +
  xlim(0,100) +
  scale_x_continuous(breaks=seq(0, 100, 20), limits=c(0, 100)) 
allsnpsurv
ggsave("snp_fit_scatter_surv.png")


snp3surv <- snp %>% 
  filter(ucc3==1 & percpop<40) %>% 
  ggplot(aes(xmin = access_survlb, xmax=access_survub, y = percpop, color = as.factor(snp), alpha = 0.5)) +
  geom_linerange() +
  geom_point(aes(x=access_surv, y=percpop, color=as.factor(snp)), alpha=0.5) +
  scale_alpha(guide = "none") +
  scale_color_brewer(palette = pall) +
  labs(color="SNP",
       title = "SNP Councils 3+ years post-campaign (n=416)",
       x = "Survey ITN access",
       y = "")  +
  xlim(0,100) +
  scale_x_continuous(breaks=seq(0, 100, 20), limits=c(0, 100)) +
  geom_smooth(aes(x = access_surv, y = percpop), se = FALSE, inherit.aes = TRUE, method = "glm", color = "steelblue") +
  theme(legend.position = "none")
ggsave("snp3_fit_scatter_surv.png")

plot_grid(allsnpsurv, snp3surv, labels = "AUTO")
ggsave("snp_bifit_scatter_surv.png")

```


```{r faceted}
snpf <- snp %>% 
  filter(ucc3==1 & outlier==0)

facets <- snpf %>% 
  ggplot(aes(x = access_est, y = percpop, color = as.factor(snp), alpha = 0.5)) +
  geom_point(shape=20) +
  scale_alpha(guide = "none") +
  scale_color_brewer(palette = pall) +
  labs(color="SNP",
       # title = paste0("SNP Councils ≥3 years post-campaign (n=",nrow(snpf),")"),
       x = "Predicted ITN access",
       y = "Nets issued as a percentage of the population (NPP)")  +
  xlim(0,100) +
   scale_y_continuous(breaks=seq(0, 40, 10), limits=c(0, 40)) +
 scale_x_continuous(breaks=seq(0, 100, 20), limits=c(0, 100)) +
  geom_smooth(data = snpf, aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "glm", span=2, color = "steelblue") +
   geom_hline(yintercept=15, alpha=0.4) +
  geom_vline(xintercept=c(60,80), alpha=0.4) +
  theme_few() +
  theme(legend.position = "none",
        # plot.title = element_text(size = 9),
        axis.text=element_text(size=6)) +
  facet_wrap(facets=~label, nrow=2)


facets
ggsave("snp_facets.png")
```


``` {r allinone}
span <- 3
lcolor <- "steelblue"
alpha <- 0.02
size=0.5

multi <- snp %>% 
  filter(ucc3==1 & outlier==0) %>% 
  ggplot(aes(x = access_est, y = percpop, color = as.factor(snp), alpha = 0.5)) +
  geom_point(shape=20) +
  scale_alpha(guide = "none") +
  scale_color_brewer(palette = pall) +
  labs(color="SNP",
       title = "SNP Councils ≥3 years post-campaign (n=416)",
       x = "Predicted ITN access",
       y = "")  +
  xlim(0,100) +
   scale_y_continuous(breaks=seq(0, 40, 10), limits=c(0, 40)) +
 scale_x_continuous(breaks=seq(0, 100, 20), limits=c(0, 100)) +
  # geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & snp==1), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "loess", span=span, color = lcolor, alpha=alpha, size=size) +
  # geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & snp==2), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "loess", span=span, color = lcolor, alpha=alpha, size=size) +
  # geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & snp==3), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "loess", span=span, color = lcolor, alpha=alpha, size=size) +
  # geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & snp==4), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "loess", span=span, color = lcolor, alpha=alpha, size=size) +
  # geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & snp==5), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "loess", span=span, color = lcolor, alpha=alpha, size=size) +
  # geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & snp==6), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "loess", span=span, color = lcolor, alpha=alpha, size=size) +
  # geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & snp==7), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "loess", span=span, color = lcolor, alpha=alpha, size=size) +
  # geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & snp==8), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "loess", span=span, color = lcolor, alpha=alpha, size=size) +
  geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & snp==9), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "loess", span=span, color = lcolor, alpha=alpha, size=size) +
  # geom_smooth(data = subset(snp, ucc3==1 & outlier==0 & access_est>40), aes(x = access_est, y = percpop), se = FALSE, inherit.aes = FALSE, method = "glm", color = "red") +
  theme(legend.position = "none",
        plot.title = element_text(size = 9)) +
  geom_hline(yintercept=15, alpha=0.4) +
  geom_vline(xintercept=c(60,80), alpha=0.4)
  # geom_segment(aes(x=0, y=15, yend=15, xend=60), color="dimgrey", alpha=0.2, size=0.1) +
  # geom_segment(aes(x=60, y=0, xend=60, yend=15), color="dimgrey", alpha=0.2, size=0.1)
```

```{r}
plot_grid(allsnp, multi, labels = "AUTO")
ggsave("snp_multifit_scatter.png", width=7, height=5, units="in")
```

```{r}
plot_grid(allsnp, facets, labels = "AUTO", rel_widths = c(1.5,2))
ggsave("snp_facet_scatter.png", width=8, height=5, units="in")
```

```{r facetsurveys}
facetsurv <- snp %>% 
  filter(ucc3==1 & percpop<40) %>% 
  ggplot(aes(xmin = access_survlb, xmax=access_survub, y = percpop, color = as.factor(snp), alpha = 0.5)) +
  # geom_linerange() +
  geom_point(aes(x=access_surv, y=percpop, color=as.factor(snp)), alpha=0.5) +
  scale_alpha(guide = "none") +
  scale_color_brewer(palette = pall) +
  labs(color="SNP",
       title = "SNP Councils 3+ years post-campaign",
       x = "Survey ITN access",
       y = "")  +
  xlim(0,100) +
  scale_x_continuous(breaks=seq(0, 100, 20), limits=c(0, 100)) +
  geom_smooth(aes(x = access_surv, y = percpop), se = FALSE, inherit.aes = TRUE, method = "loess", color = "steelblue") +
  geom_hline(yintercept=15, alpha=0.4) +
  geom_vline(xintercept=c(60,80), alpha=0.4) +
  theme_few() +
  theme(legend.position = "none",
        plot.title = element_text(size = 9),
        axis.text=element_text(size=6)) +
  facet_wrap(facets=~label, nrow=2)


facetsurv
ggsave("facet_surveys.png")

```

```{r violins}
summary_df <- snpf %>% filter(percpop<79) %>% group_by(snp) %>% summarize(m=mean(percpop))

boxes <- snpf %>% 
  filter(percpop<79) %>% 
  mutate(snpminus=snp-0.15,
         snpplus=snp+0.15) %>% 
ggplot() +
  geom_boxplot(aes(snp, percpop, group=snp, color="black"), outlier.alpha=0.1) +
  labs(x = "Round of SNP",
       y = "Percent",
       color="") +
  scale_x_continuous(breaks=seq(0,10,1), limits=c(0,10)) +
 geom_boxplot(aes(snpminus, access_est, group=snp, color="green"), width=.25, outlier.alpha=0.1) +
  geom_boxplot(aes(snpplus, access_surv, group=snp, color="blue"), width=.25, outlier.alpha=0.1) +  
  # geom_jitter(aes(snp, percpop, group=snp, color="black"), width=0.2, alpha=0.1) +
  # geom_jitter(aes(snpminus, access_est, group=snp, color="green"), width=0.2, alpha=0.1) +
  # geom_jitter(aes(snpplus, access_surv, group=snp, color="blue"), width=0.2, alpha=0.1) +
  scale_color_hue(labels=c("NPP","Survey ITN access","Predicted ITN access")) +
  theme(legend.position="bottom") +
  # geom_hline(yintercept=15) +
  geom_text(data=summary_df, aes(snp, m, label=round(m,1)), size=3, vjust=-2)

boxes
ggsave("boxes.png")
plot_grid(facets, boxes, labels="AUTO")
ggsave("facets_boxes_combo.png", width=8, height=5, units="in")

```

```{r}
ggbetweenstats(data=snp, x=snp, y=percpop)
```

```{r violinboxes}
vboxes <- snpf %>% 
  filter(percpop<79) %>% 
ggplot() +
  geom_violin(aes(snp, percpop, group=snp, color="black")) +
  labs(x = "Round of SNP",
       y = "Percent",
       color="") +
  scale_x_continuous(breaks=seq(0,10,1), limits=c(0,10)) +
 geom_violin(aes(snp, access_est, group=snp, color="green")) +
  geom_violin(aes(snp, access_surv, group=snp, color="blue"), position = position_dodge(width=.75)) +
  # geom_boxplot(aes(snp, percpop, group=snp, color="black"), notch=FALSE, varwidth=TRUE, alpha=0.2) +
  geom_jitter(aes(snp, percpop, group=snp, color="black"), width=0.2, alpha=0.1) +
  geom_jitter(aes(snp, access_est, group=snp, color="green"), width=0.2, alpha=0.1) +
  geom_jitter(aes(snp, access_surv, group=snp, color="blue"), width=0.2, alpha=0.1) +
  scale_color_hue(labels=c("NPP","Survey ITN access","Predicted ITN access")) +
  theme(legend.position="bottom") +
  # geom_hline(yintercept=15) +
  geom_text(data=summary_df, aes(snp, m, label=round(m,1)), size=3, fontface="bold")

vboxes

plot_grid(facets, vboxes, labels="AUTO")
ggsave("facets_vboxes_combo.png", width=8, height=5, units="in")
```

```{r ribbons}

lb <- 0.25
ub <- 0.75

npp <- snp %>% 
  group_by(snp) %>% 
  summarize(npp=mean(percpop), npplb=quantile(percpop, lb, prob = lb), nppub=quantile(percpop, ub, prob=ub))


pred <- snp %>% 
  group_by(snp) %>% 
  summarize(est=mean(access_est), estlb=quantile(access_est, lb, prob = lb), estub=quantile(access_est, ub, prob=ub))

sur <- snp %>% 
  group_by(snp) %>% 
  summarize(sur=mean(access_surv), surlb=quantile(access_surv, lb, prob = lb), surub=quantile(access_surv, ub, prob=ub))

collapse <- npp %>% 
  left_join(pred, by="snp") %>% 
  left_join(sur, by="snp")

collapse %>% 
  ggplot() +
  geom_ribbon(aes(x=snp, ymin=npplb, ymax=nppub, fill="NPP"), alpha=0.2) +
  geom_ribbon(aes(x=snp, ymin=estlb, ymax=estub, fill="Predicted Access"), alpha=0.2) +
  geom_ribbon(aes(x=snp, ymin=surlb, ymax=surub, fill="Survey Access"), alpha=0.2) +
  labs(fill="",
       x="Round of SNP",
       y="Percent") +
    scale_x_continuous(breaks=seq(0,10,1), limits=c(0,10))

```

